---
title: "Applied Data Science"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages
Packages we'll look at today:

- odbc / readxl / readr / dbplyr for data access
- tidyverse for data manipulation
- DataExplorer for providing of our data
- modelr / rsamples for sampling strategy
- recipes for performing feature engineering
- glmnet / glmnetUtils / h2o / FFTrees for building models
- yardstick / broom for evaluation
- rmarkdown for documentation

## Working with databases
```{r}
library(DBI)  # talk with databases, a driver
library(odbc) # allows us to talk with DBI drivers

driver = "SQL server" # prgram that allows us to talk with the database
server = "fbmcsads.database.windows.net"
database = "WideWorldImporters-Standard"
uid = "adatumadmin"
pwd = "Pa55w.rdPa55w.rd"


con <- dbConnect(odbc(),
                 driver = driver,
                 server = server,
                 database = database,
                 uid = uid,
                 pwd = pwd)
```

